Технология использования открытых данных Федеральной налоговой службы России для экономико-географического анализа

### 2024-07-08 law-firms-geography/Paper.Rmd
Minor edits, #1

[^9]: Number of people employed (in Russian). Web resource: https://fedstat.ru/indicator/34051 (accessed 24.04.2024)
[^10]: Average annual resident population (in Russian). Web resource: https://fedstat.ru/indicator/31556 (accessed 24.04.2024)

@@

The research adheres to the general framework of regional analysis and employs the methods of descriptive statistics and data visualization, including the plotting of data on a tile-grid map of Russia. In this study, the concept of legal business is approached in a relatively narrow sense and with a strong emphasis on formal considerations. The term encompasses (1) commercial legal entities or individual entrepreneurs (2) registered in Russia according to the Russian civil law (3) whose main activity is classified under code 69.10 in the All-Russia Classifier of Kinds of Economic Activity. This technique is consistent with the selection criterion for law firms used in [@Moiseeva2016]. The analysis does not include non-profit legal services providers or specialized professionals such as notaries and advocates. Due to data availability constraints, it also omits large companies that are not present in the small and medium-sized businesses registry. In essence, the legal business, as it is defined in this context, encompasses legal counseling or similar activities. The analysis is comprised of four research problems, or questions, which are enumerated below with the description of the relevant techniques used to find the answers. The term “law firm” denotes both companies (organizations) and individuals (individual entrepreneurs). The term “lawyers” in this paper refers to the employees of these firms.

@@

Fig. 1 depicts the average relative annual number of law firms by region. Several conclusions can be drawn from this visualization. First, Moscow and Saint Petersburg stand out as the only cities in Russia where the number of law firms exceeds 100 per 100,000 inhabitants. Secondly, certain regions also have a relatively high number of law firms per capita. Geographically, these regions are situated in different parts of the country, and in some cases, they form clusters. The most prominent of these groups is located in the Middle and Southern Urals with Western Siberia and consists of Sverdlovsk, Chelyabinsk, Tyumen, Omsk, and Novosibirsk regions. Another group is located in the Far East and includes Primorye territory, Sakhalin region, and, somewhat surprisingly, Kamchatka territory. Krasnodar territory, the Republic of Tatarstan, the Republic of Karelia, and Arkhangelsk region are distinct centers with a high number of law firms per capita surrounded by the regions with a more moderate number of such firms. Moscow region and its neighboring Ivanovo region are also characterized by a considerable number of law firms per capita, but it seems that their metrics may be the result of a “spillover effect” originating from Moscow city. Third, the majority of other regions have a relatively small number of law firms, with several republics of the North Caucasus and Altai and a portion of regions of the Russian North, sometimes incorporating less than ten such companies. It can be hypothesized that the number of law firms per capita is contingent upon the economic development of a region.

The relative location of regions and groups with different numbers of law firms suggests the possibility of central-periphery relations between them. For instance, Moscow City and Region, St. Petersburg City, and the Republic of Tatarstan may serve as hubs for their surrounding areas of Central, North-Western Russia, and Volga region (Povolzhye). A similar phenomenon may be observed in other parts of the country: the Ural-Siberian cluster provides legal services for Eastern Siberia and the Far East, while the Krasnodar Territory serves the Rostov and Stavropol.

@@

```{r fig-1-counts-map, fig.cap="Fig. 1. The median annual count of law firms by regions of Russia over the period from 2016 to 2021"}
counts_map
```

@@

The regional analysis revealed several spatial features of legal business in Russia. First, the regions exhibited considerable variation in the number of law firms and lawyers. Several areas with a relatively high number of lawyers per capita were identified, including the Urals and Western Siberia cluster, a few regions of the Far East, Krasnodar territory, the Republic of Tatarstan, and a number of other regions. As might be expected, the cities of Moscow and St. Petersburg are home to the largest number of law firms and lawyers. Furthermore, it is again Moscow and St. Petersburg cities that have the highest share of lawyers in their total labor force. Nevertheless, there are other regions with a relatively high role of legal businesses in their regional economies, including Sverdlovsk, Tyumen, and Novosibirsk.

>>>>

[^9]: Number of people employed (in Russian). Web resource: https://fedstat.ru/indicator/34051 (accessed 24.04.2024)
[^10]: Average annual resident population (in Russian). Web resource: https://fedstat.ru/indicator/31556 (accessed 06.07.2024)

@@

The research adheres to the general framework of regional analysis and employs the methods of descriptive statistics and data visualization, including the plotting of data on a tile-grid map of Russia. In this study, the concept of legal business is approached in a relatively narrow sense and with a strong emphasis on formal considerations. The term encompasses commercial legal entities or individual entrepreneurs registered in Russia according to the Russian civil law whose main activity is classified under code 69.10 in the All-Russia Classifier of Kinds of Economic Activity. This technique is consistent with the selection criterion for law firms used in [@Moiseeva2016]. The analysis does not include non-profit legal services providers or specialized professionals such as notaries and advocates. Due to data availability constraints, it also omits large companies that are not present in the small and medium-sized businesses registry. In essence, the legal business, as it is defined in this context, encompasses legal counseling or similar activities. The analysis is comprised of four research problems, or questions, which are enumerated below with the description of the relevant techniques used to find the answers. The term “law firm” denotes both companies (organizations) and individuals (individual entrepreneurs). The term “lawyers” in this paper refers to the employees of these firms.

@@

Fig. 1 depicts the average relative annual number of law firms by region. Several conclusions can be drawn from this visualization. First, Moscow and Saint Petersburg stand out as the only cities in Russia where the number of law firms exceeds 100 per 100,000 inhabitants. Secondly, certain regions also have a relatively high number of law firms per capita. Geographically, these regions are situated in different parts of the country, and in some cases, they form clusters, or groups. The most prominent of these groups is located in the Middle and Southern Urals with Western Siberia and consists of Sverdlovsk, Chelyabinsk, Tyumen, Omsk, and Novosibirsk regions. Another group is located in the Far East and includes Primorye territory, Sakhalin region, and, somewhat surprisingly, Kamchatka territory. Krasnodar territory, the Republic of Tatarstan, the Republic of Karelia, and Arkhangelsk region are distinct centers with a high number of law firms per capita surrounded by the regions with a more moderate number of such firms. Moscow region and its neighboring Ivanovo region are also characterized by a considerable number of law firms per capita, but it seems that their metrics may be the result of a “spillover effect” originating from Moscow city. Third, several republics of the North Caucasus and Altai and a portion of regions of the Russian North have relatively small number of law firms, sometimes incorporating less than ten such businesses. It can be hypothesized that the number of law firms per capita is contingent upon the economic development of a region.

The relative location of regions and groups with different numbers of law firms per capita suggests the possibility of central-periphery relations between them. For instance, Moscow City and Region, St. Petersburg City, and the Republic of Tatarstan may serve as hubs for their surrounding areas of Central, North-Western Russia, and Volga region (Povolzhye). A similar phenomenon may be observed in other parts of the country: the Ural-Siberian cluster provides legal services for Eastern Siberia and the Far East, while the Krasnodar Territory serves the Rostov and Stavropol.

@@

```{r fig-1-counts-map, fig.cap="Fig. 1. The median annual count of law firms per capita by regions of Russia over the period from 2016 to 2021"}
counts_map
```

@@

The regional analysis revealed several spatial features of legal business in Russia. First, the regions exhibit considerable variation in the number of law firms and lawyers. Several areas with a relatively high number of lawyers per capita are identified, including the Urals and Western Siberia cluster, a few regions of the Far East, Krasnodar territory, the Republic of Tatarstan, and a number of other regions. As might be expected, the cities of Moscow and St. Petersburg are home to the largest number of law firms and lawyers. Furthermore, it is again Moscow and St. Petersburg cities that have the highest share of lawyers in their total labor force. Nevertheless, there are other regions with a relatively high role of legal businesses in their regional economies, including Sverdlovsk, Tyumen, and Novosibirsk.
